---
title: Comparison script for SoilR model structures
author:
- address: Hans-Knoell-Str. 10, 07745 Jena, Germany
  affiliation: '1'
  corresponding: yes
  email: jbeem@bgc-jena.mpg.de
  name: Jeffrey Beem-Miller
affiliation:
- id: '1'
  institution: Department of Biogeochemical Processes, Max Planck Institute for Biogeochemistry,
    Jena, Germany
output:
  pdf_document: papaja::apa6_pdf
  html_notebook:
    css: custom.css
    toc: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
  word: papaja::apa6_docx
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
documentclass: apa6
draft: no
figsintext: yes
figurelist: yes
floatsintext: yes
footnotelist: no
always_allow_html: true
linenumbers: yes
mask: no
classoption: man
shorttitle: SoilR model structures
tablelist: no
header_includes:
- \usepackage[utf8]{inputenc}
- \usepackage{float}
- \usepackage{longtable}
- \setlength{\headheight}{28pt}
---
```{r global_options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = 'center', dev = c('cairo_pdf', 'png'),
                      crop = NULL)
options(scipen = 5)
```

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(SoilR)
library(readxl)
library(ISRaD)
library(emmeans)
library(gt)
library(GSIF)
library(aqp)
library(cowplot)
library(papaja)
library(kableExtra)
library(grid)
library(gridExtra)
library(FME)
library(forecast)
```

```{r col-pals}
# color palettes for ECO & PM
warm <- "#BF812D"
cool <- "#80CDC1"
cold <- "#01665E"
granite <- "#9daba9"
andesite <- "#382dbf"
basalt <- "#bf382d"
```

```{r load raw-data-ingest fx}
source("./utilities/jena_ams_ingest.R")
source("./utilities/jena_elm_ingest.R")
```

```{r atm-14c}
# use Hua2021 (SoilR package)
NHZone2 <- read_excel(
  "../data/external/Hua_2021/S0033822221000953sup002.xls", sheet = 2, skip = 5,
  col_names = c("Year.AD", "mean.Delta14C", "sd.Delta14C", "mean.F14C", "sd.F14C")) %>%
  data.frame

# forcast data through 2021 (2 additional years)
yrs <- seq(2000, 2019.25, by = 1/4) # A series of years by quarters
nz2 <- spline(NHZone2[ , c(1, 4)], xout = yrs) # quarterly spline interpolation of fm data
nhz2 <- ts(nz2$y, start = 2000, freq = 4) # Transformation into a time-series object
m <- ets(nhz2) # Fits an exponential smoothing state space model to the time series
f2 <- forecast(m, h = 2.75 * 4) # Uses the fitted model to forecast 2 years into the future

# bind pre and post-bomb curves; add forecasted data
atm14c <- rbind(
  bind.C14curves(IntCal20, NHZone2, "AD"),
  data.frame(Year.AD = seq(tsp(f2$mean)[1],tsp(f2$mean)[2], by=1/tsp(f2$mean)[3]),
             Delta14C = suppressWarnings(
               convert_fm_d14c(
                 fm = as.numeric(f2$mean), obs_date_y = seq(2019.75, 2022, by = .25), verbose = FALSE)),
             Sigma = NA)
)

# filter to 1900-2022 and calc annual averages
Datm <- data.frame(Date = seq(1900.5, 2021.5), d14c = NA)
for (i in seq_along(Datm$Date)) {
  ix <- which(atm14c$Year.AD >= Datm[i, "Date"] & atm14c$Year.AD < Datm[i, "Date"] + 1)
  Datm[i, "d14c"] <- mean(atm14c[ix, "Delta14C"], na.rm = TRUE)
}
```

# Notes:
1. bulk & respired 14C used for contraints
  * three time points for bulk 14C (2001, 2009, 2019)
  * two time points for respired CO2 (2001, 2019)
2. steady-state C stocks (mean of 2001, 2009 stock data) used to determine inputs

*This script requires the following input: obs.bulk.14c, obs.resp.14c, obs.Cstock*

# Workflow
1. Fit models by eye for initial parameter estimates
2. Run modfit using bulk and respired 14C time series data
3. Run bayesian parameter optimization
4. Reject models w/ stocks outside ±2SE of measured values
5. Adjust inputs w/ modeled stocks and reject models w/ unrealistic inputs
    * If inputs are not realistic [how to come up with realistic ranges?], may have to implement some constraints on parameter optimization, e.g. if (In > max | In < min) { error } else { fit pars}. Could calculate rejection rate associated with this process...
6. Compare models using AIC; other metrics? E.g. collinearity of parameters, ssr, bias, etc.
7. Get ages and transit times from "best fit" models, e.g. lowest AIC value
8. Compare distribution of ages/transit times with age distribution from density fractionation, thermal fractionation
9. Compare ages, transit times, to mineral characteristics (PCM, minC proportion, etc.)
10. Assess model structure/complexity [objective metrics here?]

# Questions
1. How to set acceptable ranges for inputs?
2. How to deal with lag times? Would be nice to avoid fitting a lag parameter. Options: could set lag by depth, but likely the optimal lag value will vary by site. Could fit lag parameter for 1p models. Would this then be relevant for more complex systems?

**Should copy relevant code for preparing obs.bulk.14c, obs.resp.14c**

```{r read-14c-data, include = FALSE}
# 2001, 2019
ams_jena_results_dirs <- list.files("../data/raw", pattern = "ams_jena_results", full.names = TRUE)
ams_results_ls <- lapply(seq_along(ams_jena_results_dirs), function(i) {
  read_jena_ams_results(ams_jena_results_dirs[i])
})
names(ams_results_ls) <- list.files("../data/raw", pattern = "ams_jena_results")

# 2009
sra.09.all <- read_excel(
  "../data/external/sra_ras_sum/sierra_data_summary_2020.xlsx", 
  sheet = "Data_Summary_2018_paper") %>% 
  type.convert(., as.is = TRUE) %>% 
  mutate(mass_kgm2 = Soil_finefraction * Thickness_cm * BD_g_cm_3 * 10) %>%
  data.frame
pro_nms <- substr(unique(sra.09.all$pro_name), 1, 4)
sra.2009.ls <- split(sra.09.all, sra.09.all$pro_name)
names(sra.2009.ls) <- pro_nms
```

# load pre-built bulk, respired 14C objs
```{r obs-blk-rsp}
load("/Users/jeff/sra-ts/source/obs.resp.14c.RData")
load("/Users/jeff/sra-ts/source/obs.bulk.14c.RData")
```

# SOC stocks
```{r obs-blk-c-dat}
# 2019
elm_results_dir <- list.files("../data/raw", pattern = "elm_jena_results-soil", full.names = TRUE)
elm_results_ls <- suppressMessages(lapply(seq_along(elm_results_dir), function(i) {
  template_file <- "../data/raw/elm_jena_template/elm_jena_template2.xls"
  read_jena_elm_results(elm_results_dir[i], template_file)
}))
names(elm_results_ls) <- list.files("../data/raw", pattern = "elm_jena_results-soil")

# extract to df
sra.19.blkC.all <- bind_rows(unlist(elm_results_ls, recursive = FALSE)) %>%
  mutate(PMeco = sapply(strsplit(ID, "_"), "[", 2),
         pro_rep = sapply(strsplit(ID, "_"), "[", 3),
         depth = sapply(strsplit(ID, "_"), "[", 4),
         year = 2019)

# 0-30 cm only
sra.19.blkC <- bind_rows(
  lapply(split(sra.19.blkC.all, sra.19.blkC.all$PMeco), function(df) {
    df$lyr_bot <- as.numeric(unlist(lapply(strsplit(df$depth, "-"), "[[", 2)))
    df <- df[df$lyr_bot < 31, ]
    df$pro_name <- paste(df$PMeco, df$pro_rep, sep = "_")
    return(df[ , c("year", "PMeco", "pro_rep", "depth", "lyr_bot", "C", "N")])
  }))
save(sra.19.blkC, file = "sra.19.blkC.RData")

# c-spline fx
sp.fx <- function(df, sp_var_in, sp_var_out, d = t(c(0, 10, 20, 30)), year) {
  out <- bind_rows(
    lapply(split(df, df$PMeco), function(x) {
      depths(x) <- pro_name ~ lyr_top + lyr_bot
      sp <- suppressMessages(
        mpspline(x, var.name = sp_var_in, d = d, show.progress = FALSE))$var.std
      sp$rep <- seq(1, nrow(sp))
      sp %>%
        pivot_longer(cols = 1:3, names_to = "depth", values_to = sp_var_out)
    }), .id = "PMeco") %>%
    mutate(lyr_bot = as.numeric(substr(sapply(strsplit(depth, "-"), "[", 2), 1, 2)),
           year = year) %>% data.frame
  out[ , c("PMeco", "rep", "year", "lyr_bot", sp_var_out)]
}

## 2001
sra.01.blkC <- read_excel("../data/external/sra_ras_sum/sierra_data_summary_2020.xlsx", sheet = "2001_bulk_data") %>% type.convert(., as.is = TRUE) %>% mutate(pro_name = paste(PMeco, pro_rep, sep = "_")) %>% data.frame 
sra.01.blkC <- sra.01.blkC[c(which(grepl("AN", sra.01.blkC$ID)), which(grepl("BS", sra.01.blkC$ID)), which(grepl("GR", sra.01.blkC$ID))), ]
sra.01.blkC <- sra.01.blkC[which(!is.na(sra.01.blkC$C.)), ]
# convert lyr_top to numeric and add pro_name col
sra.01.blkC$lyr_top <- as.numeric(sra.01.blkC$lyr_top)
sra.01.blkC$pro_name <- paste0(sra.01.blkC$PMeco, sra.01.blkC$pro_rep)
sra.01.blkC$lyr_soc <- sra.01.blkC$bd.g.cm3 * sra.01.blkC$fine.earth. * sra.01.blkC$C. * (sra.01.blkC$lyr_bot - sra.01.blkC$lyr_top) * 10^-3
blkC.rep.soc.sp.01.df <- sp.fx(sra.01.blkC, "lyr_soc", "lyr_soc", year = 2001)

## 2009
sra.09.blkC <- read_excel("../data/external/sra_ras_sum/sierra_data_summary_2020.xlsx", sheet = "2009_bulk_data") %>% type.convert(., as.is = TRUE) %>% data.frame
sra.09.blkC$PMeco <- paste0(toupper(substr(sra.09.blkC$Parent_Material, 1, 2)),
                            tolower(sra.09.blkC$Biome))
sra.09.blkC$PMeco <- ifelse(
  substr(sra.09.blkC$PMeco, 1, 2) == "BA", 
  paste0("BS", substr(sra.09.blkC$PMeco, 3, 4)), 
  sra.09.blkC$PMeco)
names(sra.09.blkC)[which(names(sra.09.blkC) == "top.mineral")] <- "lyr_top"
names(sra.09.blkC)[which(names(sra.09.blkC) == "bottom.mineral")] <- "lyr_bot"
sra.09.blkC$lyr_soc <- sra.09.blkC$Thickness_cm * sra.09.blkC$BD_g_cm_3 * sra.09.blkC$Soil_finefraction * sra.09.blkC$C_pct * 10^-1
# fit spline to C data (0-10, 10-20, 20-30 cm output)
blkC.rep.soc.sp.09.df <- sp.fx(sra.09.blkC, "lyr_soc", "lyr_soc", year = 2009)

## SOC stocks
# use mean of 2001 and 2009 SOC stocks for steady-state estimate
csoc.19.0_30.df <- rbind(blkC.rep.soc.sp.09.df, blkC.rep.soc.sp.01.df) %>%
  mutate(lyr_top = lyr_bot - 10,
         lyr_name = paste0(PMeco, "_", lyr_top, "-", lyr_bot))
csoc.19.0_30.sum.df <- csoc.19.0_30.df %>%
  group_by(PMeco, lyr_top, lyr_bot) %>%
  summarize(across(lyr_soc, .fns = list(mean = mean, sd = sd)),
            n = n(), .groups = "drop") %>%
  mutate(lyr_name = paste0(PMeco, "_", lyr_top, "-", lyr_bot),
         lyr_soc_lwr = lyr_soc_mean - lyr_soc_sd / sqrt(n) * 2,
         lyr_soc_upr = lyr_soc_mean + lyr_soc_sd / sqrt(n) * 2)
csoc.19.0_30.sum.ls <- split(csoc.19.0_30.sum.df, csoc.19.0_30.sum.df$lyr_name)
csoc.19.0_30 <- lapply(csoc.19.0_30.sum.ls, function(df) 
  data.frame(lyr_soc = df$lyr_soc_mean, lyr_soc_lwr = df$lyr_soc_lwr, lyr_soc_upr = df$lyr_soc_upr))
obs.cStock <- lapply(split(csoc.19.0_30.df, csoc.19.0_30.df$lyr_name), function(x)
      x %>% rename(time = year, cStock = lyr_soc) %>% select(time, cStock) %>% data.frame)
```

```{r mod-inputs}
# start with In = 1
in.1.ls <- lapply(vector(mode = "list", length = 27), function(x) x <- 1)
```

```{r mod-fxs}
# lambda (true half-life of 14C)
lambda <- 1 / 8267

# get initial fm value from k (pre-bomb)
fm <- function (k){
  k/(k + lambda)
}

# get k from initial fm (pre-bomb)
k <- function (fm) {
  (fm * lambda)/(1 - fm)
}

# splicing indices
ix.10 <- seq(1, 27, 3)
ix.20 <- seq(2, 27, 3)
ix.30 <- seq(3, 27, 3)

# Atm 14C for years in which bulk/resp 14C are known
atm14c.01.09.19 <- data.frame(
  year = c(2001, 2009, 2019),
  d14c = c(mean(atm14c[which(atm14c$Year.AD > 2001 & atm14c$Year.AD < 2002), "Delta14C"]),
           mean(atm14c[which(atm14c$Year.AD > 2009 & atm14c$Year.AD < 2010), "Delta14C"]),
           mean(atm14c[which(atm14c$Year.AD > 2019 & atm14c$Year.AD < 2020), "Delta14C"])))

# constraint df fx
con.df.fx <- function(PMeco_depth) {
  bulk.df <- obs.bulk.14c[[PMeco_depth]]
  resp.df <- obs.resp.14c[[PMeco_depth]]
  return(
    con.df <- data.frame(pool = c(rep("bulkC", nrow(bulk.df)), rep("respiration", nrow(resp.df))),
                         d14c = c(bulk.df$bulkC, resp.df$resp),
                         Year = c(bulk.df$time, resp.df$time)))
}

# C stocks function
soc.fx <- function(pars, mod, In, out = "pools", mod_mat = FALSE) {

  # steady-state stock calc fx
  calc.soc <- function(A, in_vector) {
    (-1 * solve(A) %*% in_vector)
  }
  
  # model matrices
  if (mod == "1p") {
    
    # 1p mod matrix
    A <- pars
    
    # 1p steady-state stocks
    in_vector <- In
    ss.cstock <- pars^-1 * in_vector
    
    # pool names
    pnms <- "bulkC"
    
  } else if (grepl("2p", mod)) {
    
    # pool names
    pnms <- c("fast", "slow")
    
    # 2pp mod matrix
    A <- -diag(pars[1:2])
    
    # 2pp steady-state C stocks
    in_vector <- c(In * pars[3], In * (1 - pars[3]))
    ss.cstock <- calc.soc(A, in_vector)
    
    if (mod == "2ps") {
      
      # 2ps mod matrix
      A[2, 1] <- pars[3] * pars[1]
      
      # 2ps steady-state C stocks
      in_vector <- c(In, 0)
      ss.cstock <- calc.soc(A, in_vector)
    }
  } else if (grepl("3p", mod)) {
    
    # pool names
    pnms <- c("fast", "slow", "passive")
    
    # 3pp mod matrix
    A <- -diag(pars[1:3])
    
    # 3pp steady-state C stocks
    in_vector <- c(In * pars[4], In * pars[5], In * (1 - pars[4] - pars[5]))
    ss.cstock <- calc.soc(A, in_vector)
    
    if (mod == "3ps") {
      
      # 3ps mod matrix
      A[2, 1] <- pars[4] * pars[1]
      A[3, 2] <- pars[5] * pars[2] 
      
      # 3ps steady-state C stocks
      in_vector <- c(In, 0, 0)
      ss.cstock <- calc.soc(A, in_vector)
    }
  }
  
  if (out == "sum") {
    soc <- sum(ss.cstock)
    if (mod_mat) {
      list(A_mat = A, in_vector = in_vector, ss_soc = soc, par_names = pnms)
    } else {
      soc
    }
  } else {
    soc <- ss.cstock
    if (mod_mat) {
      list(A_mat = A, in_vector = in_vector, ss_soc = soc, par_names = pnms)
    } else {
      soc
    }
  }
}

# model fx
modFun <- function(pars, mod, In, lag = 0, pass = TRUE, verbose = TRUE, out = "modFit", PMeco_depth = NULL, var_14c = "d14c") {
  
  # run soc.fx to get: mod_mat [[1]], in_vector [[2]], steady-state C [[3]], and pool names [[4]]
  soc.fx_out <- soc.fx(pars, mod, In, mod_mat = TRUE)
  ss.cstock <- soc.fx_out[[3]]
  
  # check for negative stocks
  if (any(ss.cstock <= 0)) {
    cat("pool ", which(ss.cstock <= 0), "< 0")
  }
  
  # time index
  ix.t <- c((lag + 1):nrow(Datm))
  
  # check for 1p mod
  if (mod == "1p") {
    model <- OnepModel14(
      t = Datm$Date[ix.t],
      k = soc.fx_out[[1]],
      C0 = as.vector(ss.cstock),
      F0_Delta14C = Delta14C_from_AbsoluteFractionModern(fm(soc.fx_out[[1]])),
      In = soc.fx_out[[2]],
      inputFc = Datm,
      lag = lag,
      pass = pass)
  } else {
    
    # calculate initial 14C
    F0_Delta14C <- unlist(
      lapply(-diag(soc.fx_out[[1]]), function(x) Delta14C_from_AbsoluteFractionModern(fm(x))))
    
    # multipool model fx
    mod.fx <- function(A,
                       t,
                       in_vector,
                       C0,
                       F0_Delta14C, 
                       xi = 1, # timestep
                       inputFc, 
                       lag = lag,
                       pass = pass) {
      t_start = min(t)
      t_stop = max(t)
      inputFluxes = BoundInFluxes(function(t) {
          matrix(nrow = length(in_vector), ncol = 1, in_vector)
        }, t_start, t_stop)
      if (length(xi) == 1) 
        fX = function(t) {
          xi
        }
      At = BoundLinDecompOp(map = function(t) {
        fX(t) * A
      }, t_start, t_stop)
      Fc = BoundFc(inputFc, lag = lag, format = "Delta14C")
      mod = Model_14(t, At, ivList = C0, initialValF = ConstFc(F0_Delta14C, "Delta14C"), 
                     inputFluxes = inputFluxes, inputFc = Fc, pass = pass)
    }
    
    # run model
     model <- mod.fx(
      A = soc.fx_out[[1]],
      t = Datm$Date[ix.t],
      in_vector = soc.fx_out[[2]],
      C0 = as.vector(ss.cstock),
      F0_Delta14C = F0_Delta14C,
      inputFc = Datm,
      lag = lag,
      pass = pass) 
    }

  # get mod values
  C14m <- getF14C(model)
  C14p <- getF14(model) 
  C14r <- getF14R(model)
  Ctot <- getC(model)
  
  if (var_14c == "fm") {
    dates <- Datm$Date[ix.t]
    for (i in seq_along(dates)) {
      C14m[i] <- convert_fm_d14c(d14c = C14m[i], obs_date_y = dates[i], verbose = FALSE) 
      C14p[i] <- convert_fm_d14c(d14c = C14p[i], obs_date_y = dates[i], verbose = FALSE) 
      C14r[i] <- convert_fm_d14c(d14c = C14r[i], obs_date_y = dates[i], verbose = FALSE)
    }
  }
  
  if (out == "modFit") {
    # dataframe for modFit fx
    data.frame(
      time = Datm$Date[ix.t],
      resp = C14r,
      bulkC = rowSums(Ctot[1, ] / sum(Ctot[1, ]) * C14p),
      cStock = sum(Ctot[1, ]))
  } else {
    
    # sum c stocks
    ss.cstock <- round(ss.cstock, 2)
    cstock.sum <- ifelse(is.null(dim(ss.cstock)), ss.cstock, colSums(ss.cstock))
    
    if (verbose) {
      # print site and steady-state stocks
      if (!is.null(PMeco_depth)) cat(paste0(PMeco_depth, "\n"))
      for (i in seq_along(ss.cstock)) {
        cat(paste(soc.fx_out[[4]][i], ss.cstock[i], "\n"))
      }
      cat(cstock.sum, " (modeled total C stock)\n")
      if (!is.null(PMeco_depth)) {
        cat(round(csoc.19.0_30[[PMeco_depth]][ , "lyr_soc"], 1), " (measured total C stock)\n")
        if (cstock.sum < csoc.19.0_30[[PMeco_depth]][ , "lyr_soc_lwr"]) cat("Modeled stocks too low w/ current inputs\n") else
          if (cstock.sum > csoc.19.0_30[[PMeco_depth]][ , "lyr_soc_upr"]) cat("Modeled stocks too high w/ current inputs\n")
      }
    }
    
    # data frame for plotting
    data.frame(
      years = rep(Datm$Date[ix.t], (ncol(C14p) + 3)),
      d14C = c(c(C14p),
               C14m,
               C14r,
               Datm$d14c[ix.t]),
      pool = rep(c(soc.fx_out[[4]], "bulkC", "respiration", "atm"), 
                 each = nrow(C14p))) %>%
      distinct
  }
}
```

```{r mod-plot-fx}
C14.plot.fx <- function(plot.df, con.df, mod, pool_filter = NULL, PMeco_depth) {
    
  # set line and color values
  cvals <- c("atm" = 8, "fast" = "#e41f88", "slow" = "#1f88e4", "passive" = "#1fe47b", "respiration" = "#e47b1f", "bulkC" = "black")
  lvals <- c("atm" = 1, "fast" = 2, "slow" = 2, "passive" = 2, "respiration" = 1, "bulkC" = 1)
  
  # filter pools as desired
  if (!is.null(pool_filter)) {
    plot.df <- plot.df[which(plot.df$pool %in% pool_filter), ]
  }
  
  # plot
  plot.df %>%
    ggplot(., aes(years, d14C, color = pool, linetype = pool)) +
    geom_path() +
    geom_point(data = con.df, aes(Year, d14c, color = pool), size = 3) +
    scale_color_manual(name = "Pool", values = cvals) +
    scale_linetype_manual(values = lvals) +
    scale_x_continuous(limits = c(1950, 2022)) +
    ggtitle(paste(PMeco_depth, mod)) +
    xlab("Year") +
    ylab(expression(''*Delta*''^14*'C (‰)')) +
    guides(linetype = "none") +
    theme_bw() +
    theme(panel.grid = element_blank())
}
```

# Load initial parameter lists (1p, 2pp, 2ps, 3pp, 3ps)
```{r pars-i-1p}
load("pars.i.1p.RData")
load("pars.i.2pp.RData")
load("pars.i.2ps.RData")
```

# Optimize parameter sets
```{r modFit-fx-multipool}
mod.fits.fx <- function(pars, mod, In, sub, lag = 0, upper, lower, cost, var_14c = "d14c") {
  
  if (var_14c == "d14c") {
    obs.bulk <- obs.bulk.14c
    obs.resp <- obs.resp.14c
  } else {
    obs.bulk <- obs.bulk.fm
    obs.resp <- obs.resp.fm
  }
  
  # start loop
  lapply(seq_along(pars[sub]), function(i) {
    
    # start timer and print PMeco_depth
    start <- Sys.time()
    cat(paste0(names(pars)[sub][i], " parameter fitting\n"))
    
    # set vars
    PMeco_depth <- names(pars)[sub][i]
    pars <- pars[sub][[i]]
    
    # Set input
    In <- In[sub][[i]]
    
    # define cost function
    if (cost == "14C + cStock") {
      mod.Cost <- function(pars) {
        modelOutput <- modFun(pars, mod, In, lag, verbose = FALSE, var_14c = var_14c)
        cost1 <- modCost(model = modelOutput, obs = obs.bulk[sub][[i]], scaleVar = TRUE)
        cost2 <- modCost(model = modelOutput, obs = obs.resp[sub][[i]], scaleVar = TRUE, cost = cost1) 
        return(modCost(model = modelOutput, obs = obs.cStock[sub][[i]], cost = cost2))
      }
    } else if (cost == "14C") {
      mod.Cost <- function(pars) {
        modelOutput <- modFun(pars, mod, In, lag, verbose = FALSE, var_14c = var_14c)
        cost1 <- modCost(model = modelOutput, obs = obs.bulk[sub][[i]], scaleVar = TRUE)
        return(modCost(model = modelOutput, obs = obs.resp[sub][[i]], scaleVar = TRUE, cost = cost1))
      } 
    }
    
    # fit pars
    fit <- tryCatch(
      modFit(f = mod.Cost,
             p = pars,
             method = 'Nelder-Mead',
             lower = lower,
             upper = upper),
      error = function (e) {cat("ERROR :", conditionMessage(e), "\n")})      
      
    # parameter sensitivity
    Sfun <- sensFun(mod.Cost, fit$par)
    
    # End timer and print elapsed time
    end <- Sys.time()
    cat(paste0("time: ", end - start, "\n"))
    
    # Return fitted parameters and sensitivity
    return(list(modfit = fit, sens = Sfun))
  }) 
}
```

```{r modFit-fx-1pool}
mod.fits.1p.fx <- function(par.ls, var_14c = "d14c") {
  
  # set constraint variable
  if (var_14c == "d14c") {
    obs.bulk <- obs.bulk.14c
    obs.resp <- obs.resp.14c
  } else {
    obs.bulk <- obs.bulk.fm
    obs.resp <- obs.resp.fm
  }
  
  # start loop
  lapply(seq_along(par.ls), function(i) {

  # start timer and print PMeco_depth
  start <- Sys.time()
  cat(paste0(names(pars)[i], " parameter fitting\n"))
  
  # get k
  k <- par.ls[[i]]

  # model fx
  modFun_1p <- function(k, var_14c = var_14c) {
    mod <- OnepModel14(
      t = Datm$Date,
      k = k,
      C0 = 1 / k,
      F0_Delta14C = Delta14C_from_AbsoluteFractionModern(fm(k)),
      In = 1,
      inputFc = Datm, 
      lag = 0,
      pass = FALSE)
  C14 <- getF14C(mod)
  if (var_14c == "fm") {
    for (i in seq_along(Datm$Date)) 
      C14[i] <- convert_fm_d14c(d14c = C14[i], obs_date_y = Datm$Date[i], verbose = FALSE)
  }
  data.frame(
    time = Datm$Date,
    bulkC = C14,
    resp = C14
    )
  }
  
  # define cost fx
  mod.Cost <- function(k) {
        modelOutput <- modFun_1p(k, var_14c = var_14c)
        cost1 <- modCost(model = modelOutput, obs = obs.bulk[[i]], scaleVar = TRUE)
        return(modCost(model = modelOutput, obs = obs.resp[[i]], scaleVar = TRUE, cost = cost1)$model)
  } 
  
  # return optimization
  opt <- optim(par = k,
               fn = mod.Cost,
               method = "Brent",
               lower = 0,
               upper = 1,
               hessian = TRUE)
  
  # rename "value" to ssr
  names(opt)[which(names(opt) == "value")] <- "ssr"
  
  # retrieve variance and residuals from mod.Cost
  modelOutput <- modFun_1p(opt$par, var_14c = var_14c)
  cost1 <- modCost(model = modelOutput, obs = obs.bulk[[i]], scaleVar = TRUE)
  mod.Cost.var <- modCost(model = modelOutput, obs = obs.resp[[i]], scaleVar = TRUE, cost = cost1)$var
  mod.Cost.res <- modCost(model = modelOutput, obs = obs.resp[[i]], scaleVar = TRUE, cost = cost1)$residuals
  
  # add to opt and return
  opt$var_ms <- mod.Cost.var$SSR * mod.Cost.var$scale
  names(opt$var_ms) <- c("resp", "bulkC")
  opt$obs.mod <- mod.Cost.res[ , c("obs", "mod", "name")]
  opt$df.residual <- nrow(mod.Cost.res) - length(k) - 1
  opt$residuals <- mod.Cost.res$res
  names(opt$residuals) <- mod.Cost.res$name
    
  # End timer and print elapsed time
  end <- Sys.time()
  cat(paste0("time: ", end - start, "\n"))
  
  return(opt)
  })
}
```

```{r opt-1p-mod}
# 1p fits
mod.fits.1p <- mod.fits.1p.fx(pars.i.1p)
names(mod.fits.1p) <- names(pars.i.1p)
save(mod.fits.1p, file = paste0("../data/derived/modFit_pars/", "mod.fits.1p", "_", Sys.Date(), ".RData"))

# 2pp fit
mod.sens.fits.2pp <- mod.fits.fx(
  pars = pars.i.2pp, 
  mod = "2pp",
  In = in.1.ls,
  sub = 1:27,
  lag = 0,
  upper = c(1, 1, 1),
  lower = c(0, 0, 0),
  cost = "14C"
)
names(mod.sens.fits.2pp) <- names(pars.i.2pp)
save(mod.sens.fits.2pp, file = paste0("../data/derived/modFit_pars/", "mod.sens.fits.2pp", "_", Sys.Date(), ".RData"))


# test long data format
obs.bulk.14c.l <- lapply(obs.bulk.14c, function(x) {
  x %>% 
    group_by(time) %>% 
    summarize(across(bulkC, .fns = list(mean = mean, sd = sd)), n = n()) %>%
    mutate(name = "bulkC",
           err = bulkC_sd / sqrt(n)) %>%
    mutate(err = ifelse(is.na(err), mean(err, na.rm = TRUE), err)) %>%
    rename(bulkC = bulkC_mean) %>%
    select(name, time, bulkC, err) %>%
    data.frame
})
obs.resp.14c.l <- lapply(obs.resp.14c, function(x) {
  x %>% 
    group_by(time) %>% 
    summarize(across(resp, .fns = list(mean = mean, sd = sd)), n = n()) %>%
    mutate(name = "resp",
           err = resp_sd / sqrt(n)) %>%
    mutate(err = ifelse(is.na(err), mean(err, na.rm = TRUE), err)) %>%
    rename(resp = resp_mean) %>%
    select(name, time, resp, err) %>%
    data.frame
})

# test fraction modern as constraint
obs.bulk.fm <- lapply(obs.bulk.14c, function(x) {
  for (i in 1:nrow(x)) {
    x$bulkC[i] <- convert_fm_d14c(d14c = x$bulkC[i], obs_date_y = x$time[i], verbose = FALSE) 
  }
  return(x)
})
obs.resp.fm <- lapply(obs.resp.14c, function(x) {
  for (i in 1:nrow(x)) {
    x$resp[i] <- convert_fm_d14c(d14c = x$resp[i], obs_date_y = x$time[i], verbose = FALSE) 
  }
  return(x)
})

# std cost fx w/ additional optional arguments
mod.Cost.14c <- function(pars, mod, In = 1, lag = 0, sub = ix.10, i, verbose = FALSE, scaleVar = TRUE, wt = "none", var_14c = "d14c") {
  
  if (var_14c == "d14c") {
    obs.bulk <- obs.bulk.14c
    obs.resp <- obs.resp.14c
  } else {
    obs.bulk <- obs.bulk.fm
    obs.resp <- obs.resp.fm
  }
        modelOutput <- modFun(pars, mod, In, lag, verbose = FALSE, var_14c = var_14c)
        cost1 <- modCost(model = modelOutput, obs = obs.bulk[sub][[i]], weight = wt, scaleVar = scaleVar)
        return(modCost(model = modelOutput, obs = obs.resp[sub][[i]], weight = wt, scaleVar = scaleVar, cost = cost1))
}

# long format cost fx w/ additional optional arguments
mod.Cost.14c2 <- function(pars, mod, In = 1, lag = 0, sub = ix.10, i, verbose = FALSE, scaleVar = TRUE, var_14c = "d14c") {
        modelOutput <- modFun(pars, mod, In, lag, verbose = FALSE, var_14c = var_14c)
        cost1 <- modCost(model = modelOutput, obs = obs.bulk.14c.l[sub][[i]], y = "bulkC", err = "err", scaleVar = scaleVar)
        return(modCost(model = modelOutput, obs = obs.resp.14c.l[sub][[i]], y = "resp", err = "err", scaleVar = scaleVar, cost = cost1))
}

# costs for initial pars, standard cost fx
ANpp.10.2pp.modCost <- mod.Cost.14c(pars.i.2pp$`ANpp_0-10`, mod = "2pp", i = 1)
# costs for fitted pars, standard cost fx
ANpp.10.2pp.modCost.f <- mod.Cost.14c(ANpp.10.2pp.modFit$par, mod = "2pp", i = 1)

# costs for initial pars, fm for costs
ANpp.10.2pp.modCost.fm <- mod.Cost.14c(pars.i.2pp$`ANpp_0-10`, mod = "2pp", i = 1, var_14c = "fm")
# costs for fitted pars, standard cost fx
ANpp.10.2pp.modCost.fm.f <- mod.Cost.14c(ANpp.10.2pp.modFit$par, mod = "2pp", i = 1, var_14c = "fm")

# costs for initial pars, standard cost fx w/ errors weighted by reciprocal of sd
ANpp.10.2pp.modCost.w <- mod.Cost.14c(pars.i.2pp$`ANpp_0-10`, mod = "2pp", i = 1, wt = "std", scaleVar = FALSE)
# costs for fitted pars, standard cost fx
ANpp.10.2pp.modCost.w.f <- mod.Cost.14c(ANpp.10.2pp.modFit.w$par, mod = "2pp", wt = "std", i = 1, scaleVar = FALSE)

# costs for initial pars, long format cost fx, scaleVar = T
ANpp.10.2pp.modCost2 <- mod.Cost.14c2(pars.i.2pp$`ANpp_0-10`, mod = "2pp", i = 1, scaleVar = TRUE)
# costs for fitted pars, long format cost fx, scaleVar = T
ANpp.10.2pp.modCost2.f <- mod.Cost.14c2(ANpp.10.2pp.modFit2$par, mod = "2pp", i = 1, scaleVar = TRUE)

# costs for initial pars, long format cost fx, scaleVar = F
ANpp.10.2pp.modCost2.1 <- mod.Cost.14c2(pars.i.2pp$`ANpp_0-10`, mod = "2pp", i = 1, scaleVar = FALSE)
# costs for fitted pars, long format cost fx, scaleVar = T
ANpp.10.2pp.modCost2.1.f <- mod.Cost.14c2(ANpp.10.2pp.modFit2.1$par, mod = "2pp", i = 1, scaleVar = FALSE)

# fit model w/ initial pars and std cost fx
ANpp.10.2pp.modFit <- modFit(f = mod.Cost.14c,
                             p = pars.i.2pp$`ANpp_0-10`,
                             mod = "2pp",
                             sub = ix.10, 
                             i = 1,
                             method = 'Nelder-Mead',
                             lower = c(0, 0, 0),
                             upper = c(1, 1, 1))
ANpp.10.2pp.modFitted <- modFun(pars = ANpp.10.2pp.modFit$par, 
                                mod = "2pp",
                                In = 1,
                                out = "")
C14.plot.fx(ANpp.10.2pp.modFitted, con.df.fx("ANpp_0-10"), mod = "2pp", PMeco_depth = "ANpp_0-10")

# fit model w/ initial pars and fm cost fx
ANpp.10.2pp.modFit.fm <- modFit(f = mod.Cost.14c,
                                p = pars.i.2pp$`ANpp_0-10`,
                                var_14c = "fm",
                                mod = "2pp",
                                sub = ix.10, 
                                i = 1,
                                method = 'Nelder-Mead',
                                lower = c(0, 0, 0),
                                upper = c(1, 1, 1))
ANpp.10.2pp.modFitted.fm <- modFun(pars = ANpp.10.2pp.modFit.fm$par, 
                                   mod = "2pp",
                                   In = 1,
                                   out = "")
C14.plot.fx(ANpp.10.2pp.modFitted.fm, con.df.fx("ANpp_0-10"), mod = "2pp", PMeco_depth = "ANpp_0-10")

# fit model w/ initial pars and fm cost fx + err weighting
ANpp.10.2pp.modFit.fm.w <- modFit(f = mod.Cost.14c,
                                  p = pars.i.2pp$`ANpp_0-10`,
                                  var_14c = "fm",
                                  mod = "2pp",
                                  wt = "std",
                                  sub = ix.10, 
                                  i = 1,
                                  method = 'Nelder-Mead',
                                  lower = c(0, 0, 0),
                                  upper = c(1, 1, 1))
ANpp.10.2pp.modFitted.fm.w <- modFun(pars = ANpp.10.2pp.modFit.fm.w$par, 
                                     mod = "2pp",
                                     In = 1,
                                     out = "")
C14.plot.fx(ANpp.10.2pp.modFitted.fm.w, con.df.fx("ANpp_0-10"), mod = "2pp", PMeco_depth = "ANpp_0-10")

# fit mod w/ initial pars and std cost fx + std weighting
ANpp.10.2pp.modFit.w <- modFit(f = mod.Cost.14c,
                              p = pars.i.2pp$`ANpp_0-10`,
                              wt = "std",
                              mod = "2pp",
                              sub = ix.10, 
                              i = 1,
                              method = 'Nelder-Mead',
                              lower = c(0, 0, 0),
                              upper = c(1, 1, 1))
ANpp.10.2pp.modFitted.w <- modFun(pars = ANpp.10.2pp.modFit.w$par, 
                                mod = "2pp",
                                In = 1,
                                out = "")

# fit mod w/ initial pars and long format cost fx
ANpp.10.2pp.modFit2 <- modFit(f = mod.Cost.14c2,
                              p = pars.i.2pp$`ANpp_0-10`,
                              mod = "2pp",
                              sub = ix.10, 
                              i = 1,
                              method = 'Nelder-Mead',
                              lower = c(0, 0, 0),
                              upper = c(1, 1, 1))
ANpp.10.2pp.modFitted2 <- modFun(pars = ANpp.10.2pp.modFit2$par, 
                                mod = "2pp",
                                In = 1,
                                out = "")


ANpp.10.2pp.modFit2.1 <- modFit(f = mod.Cost.14c2,
                              p = pars.i.2pp$`ANpp_0-10`,
                              mod = "2pp",
                              sub = ix.10, 
                              i = 1,
                              method = 'Nelder-Mead',
                              lower = c(0, 0, 0),
                              upper = c(1, 1, 1))
ANpp.10.2pp.modFitted2.1 <- modFun(pars = ANpp.10.2pp.modFit2.1$par, 
                                mod = "2pp",
                                In = 1,
                                out = "")
C14.plot.fx(ANpp.10.2pp.modFitted, con.df.fx("ANpp_0-10"), mod = "2pp", PMeco_depth = "ANpp_0-10")
C14.plot.fx(ANpp.10.2pp.modFitted.w, con.df.fx("ANpp_0-10"), mod = "2pp", PMeco_depth = "ANpp_0-10")
C14.plot.fx(ANpp.10.2pp.modFitted2, con.df.fx("ANpp_0-10"), mod = "2pp", PMeco_depth = "ANpp_0-10")
C14.plot.fx(ANpp.10.2pp.modFitted2.1, con.df.fx("ANpp_0-10"), mod = "2pp", PMeco_depth = "ANpp_0-10")
C14.plot.fx(mod.fitted.2pp.i$`ANpp_0-10`, con.df.fx("ANpp_0-10"), mod = "2pp", PMeco_depth = "ANpp_0-10")
C14.plot.fx(mod.fitted.2pp$`ANpp_0-10`, con.df.fx("ANpp_0-10"), mod = "2pp", PMeco_depth = "ANpp_0-10")


# 2ps fit
mod.sens.fits.2ps <- mod.fits.fx(
  pars = pars.i.2ps, 
  mod = "2ps",
  In = in.1.ls,
  sub = 1:27,
  lag = 0,
  upper = c(1, 1, 1),
  lower = c(0, 0, 0),
  cost = "14C"
)
names(mod.sens.fits.2ps) <- names(pars.i.2ps)
save(mod.sens.fits.2ps, file = paste0("../data/derived/modFit_pars/", "mod.sens.fits.2ps", "_", Sys.Date(), ".RData"))
```

```{r }
# load fits as needed
if (!exists("mod.fits.1p")) {
 load("../data/derived/modFit_pars/mod.sens.fits.1p_2022-08-11.Rdata") 
}
if (!exists("mod.sens.fits.2pp")) {
 load("../data/derived/modFit_pars/mod.sens.fits.2pp_2022-08-11.Rdata") 
}
if (!exists("mod.sens.fits.2ps")) {
 load("../data/derived/modFit_pars/mod.sens.fits.2ps_2022-08-11.Rdata") 
}

# get mod fit only
mod.fits.2pp <- lapply(mod.sens.fits.2pp, function(x) x[[1]])
mod.fits.2ps <- lapply(mod.sens.fits.2ps, function(x) x[[1]])

# get pars
pars.fit.1p <- lapply(mod.fits.1p, "[[", 1)
pars.fit.2pp <- lapply(mod.fits.2pp, "[[", 1)
names(pars.fit.2pp) <- names(pars.i.2pp)
pars.fit.2ps <- lapply(mod.fits.2ps, "[[", 1)
names(pars.fit.2ps) <- names(pars.i.2ps)

# run fitted models
# 1p
mod.fitted.1p <- lapply(seq_along(pars.fit.1p), function(i) {
  modFun(pars = pars.fit.1p[[i]], In = in.1.ls[[i]], mod = "1p", verbose = FALSE, out = "plot.df")
})
names(mod.fitted.1p) <- names(pars.fit.1p)
# 2pp
mod.fitted.2pp.i <- lapply(seq_along(pars.i.2pp), function(i) {
  modFun(pars = pars.i.2pp[[i]], In = in.1.ls[[i]], mod = "2pp", 
         verbose = FALSE, out = "plot.df")
})
names(mod.fitted.2pp.i) <- names(pars.i.2pp)
mod.fitted.2pp <- lapply(seq_along(pars.fit.2pp), function(i) {
  modFun(pars = pars.fit.2pp[[i]], In = in.1.ls[[i]], mod = "2pp", 
         verbose = FALSE, out = "plot.df")
})
names(mod.fitted.2pp) <- names(pars.fit.2pp)
# 2ps
mod.fitted.2ps.i <- lapply(seq_along(pars.i.2ps), function(i) {
  modFun(pars = pars.i.2ps[[i]], In = in.1.ls[[i]], mod = "2ps", 
         verbose = FALSE, out = "plot.df")
})
names(mod.fitted.2ps.i) <- names(pars.i.2ps)
mod.fitted.2ps <- lapply(seq_along(pars.fit.2ps), function(i) {
  modFun(pars = pars.fit.2ps[[i]], In = in.1.ls[[i]], mod = "2ps", 
         verbose = FALSE, out = "plot.df")
})
names(mod.fitted.2ps) <- names(pars.fit.2ps)

# get ssr
get.ssr.fx <- function(mod.fit.ls, mod) {
  data.frame(bind_rows(lapply(mod.fit.ls, "[", "ssr"), .id = "PMeco_depth")) %>%
  mutate(mod = mod,
         depth = sapply(strsplit(PMeco_depth, "_"), "[[", 2))
}
ssr.1p.df <- get.ssr.fx(mod.fits.1p, "1p")
ssr.2pp.df <- get.ssr.fx(mod.fits.2pp, "2pp")
ssr.2ps.df <- get.ssr.fx(mod.fits.2ps, "2ps")
 
# combine ssr & plot
ssr.df <- rbind(
  ssr.1p.df,
  ssr.2pp.df,
  ssr.2ps.df) %>% 
  mutate(PMeco = substr(PMeco_depth, 1, 4))
ssr.df %>%
  filter(depth == "0-10") %>%
  ggplot(., aes(PMeco, ssr, fill = mod)) + 
  geom_col(position = "dodge") + 
  # facet_grid(rows = vars(depth), scales = "free") + 
  theme_bw() + 
  theme(panel.grid.minor = element_blank())

# mean residuals, by var (var_ms)
var_ms.df.fx <- function(mod.fits.ls, costs, mod) {
  df <- data.frame(bind_rows(lapply(mod.fits.ls, "[", "var_ms"), .id = "PMeco_depth"))
  df$var <- rep(costs, nrow(df)/length(costs))
  df$var_ms <- round(df$var_ms, 5)
  df$depth <- sapply(strsplit(df$PMeco_depth, "_"), "[[", 2)
  df$mod <- mod
  return(df)
}
var_ms.1p.df <- var_ms.df.fx(mod.fits.1p, c("resp", "bulkC"), "1p")
var_ms.2pp.df <- var_ms.df.fx(mod.fits.2pp, c("resp", "bulkC"), "2pp")
var_ms.2ps.df <- var_ms.df.fx(mod.fits.2ps, c("resp", "bulkC"), "2ps")

# combine var & plot
var.df <- rbind(
  var_ms.1p.df,
  var_ms.2pp.df,
  var_ms.2ps.df) %>% 
  mutate(PMeco = substr(PMeco_depth, 1, 4))
var.df %>%
  filter(depth == "0-10") %>%
  ggplot(., aes(PMeco, var_ms, fill = mod)) + 
  geom_col(position = "dodge") + 
  facet_grid(rows = vars(var), scales = "free") + 
  theme_bw() + 
  theme(panel.grid.minor = element_blank())

# summarize pars
pars.fit.1p.sum <- lapply(mod.fits.1p, function(x) {
  ## inspired by summary.modFit (FME)
  summary.1p.modFit <- function (object, cov = TRUE,...) {
    param  <- object$par
    pnames <- names(param)
    p      <- length(param)
    covar  <- try(solve(0.5*object$hessian), silent = TRUE)   # unscaled covariance
    if (!is.numeric(covar)) {
      message <- "Cannot estimate covariance; system is singular"
      warning(message)
      covar <- matrix(data = NA, nrow = p, ncol = p)
    } else message <- "ok"
  
    rownames(covar) <- colnames(covar) <-pnames
    rdf    <- object$df.residual
    resvar <- object$ssr / rdf
    se     <- sqrt(diag(covar) * resvar)
    names(se) <- pnames
    tval      <- param / se
    modVariance <- object$ssr / length(object$residuals)
  
    param <- cbind(param, se, tval, 2 * pt(abs(tval), rdf, lower.tail = FALSE))
    dimnames(param) <- list(pnames, c("Estimate", "Std. Error",
                                      "t value", "Pr(>|t|)"))
    if (cov)
      ans <- list(residuals = object$residuals,
                  residualVariance = resvar,
                  sigma = sqrt(resvar),
                  modVariance = modVariance,
                  df = c(p, rdf), cov.unscaled = covar,
                  cov.scaled = covar * resvar,
                  par = param)
    else
      ans <- list(residuals = object$residuals,
                  residualVariance = resvar,
                  sigma = sqrt(resvar),
                  modVariance = modVariance,
                  df = c(p, rdf),
                  info = object$info, niter = object$iterations,
                  stopmess = message,
                  par = param)
    class(ans) <- "summary.modFit"
    ans
  }
  summary.1p.modFit(x)
})
names(pars.fit.1p.sum) <- names(pars.fit.1p)

pars.fit.2pp.sum <- lapply(mod.fits.2pp, function(x) {
  tryCatch(summary(x), 
           error = function (e) {cat("ERROR :", conditionMessage(e), "\n")})
})
names(pars.fit.2pp.sum) <- names(pars.fit.2pp)

pars.fit.2ps.sum <- lapply(mod.fits.2ps, function(x) {
  tryCatch(summary(x), 
           error = function (e) {cat("ERROR :", conditionMessage(e), "\n")})
})
names(pars.fit.2ps.sum) <- names(pars.fit.2ps)

# compare par fits
par.fit.df.fx <- function(pars_fit, pars_fit_alt, par_set_nms) {
  df <- bind_rows(
    lapply(
      mapply(rbind, 
             pars_fit,
             pars_fit_alt,
             SIMPLIFY = FALSE), 
      function(df) {
        df <- data.frame(df)
        colnames(df) <- c("kf", "ks", "par3")
        df$est <- par_set_nms
        return(df)
      })
  )
  df$PMeco_depth <- rep(names(pars_fit), each = 2)
  df$PM <- substr(df$PMeco_depth, start = 1, stop = 2)
  df$eco <- substr(df$PMeco_depth, start = 3, stop = 4)
  df$depth <- substr(df$PMeco_depth, start = 6, stop = length(df$PMeco_depth))
  return(df)
}

# run fx for 2p mods
pars.fit.2pp.2ps.df <- par.fit.df.fx(pars_fit = pars.fit.2pp,
                                     pars_fit_alt = pars.fit.2ps,
                                     par_set_nms = c("2pp", "2ps")) 
pars.fit.2pp.2ps.PMsum.df <- pars.fit.2pp.2ps.df %>%
    group_by(PM, est) %>%
    summarize(across(.cols = c(kf, ks, par3), .fns = list(mean = mean, sd = sd)), .groups = "drop") %>%
    mutate_if(is.numeric, format, digits = 2)
pars.fit.2pp.2ps.ECOsum.df <- pars.fit.2pp.2ps.df %>%
    group_by(eco, est) %>%
    summarize(across(.cols = c(kf, ks, par3), .fns = list(mean = mean, sd = sd)), .groups = "drop") %>%
    mutate_if(is.numeric, format, digits = 2)
```

```{r mod-fit-plot-fxs}
singleMod.fit.plot.fx <- function(modFit.ls, mod) {
  lapply(seq_along(modFit.ls), function(i) {
    PMeco_depth <- names(modFit.ls)[i]
    con.df <- con.df.fx(PMeco_depth)
    C14.plot.fx(modFit.ls[[i]], con.df, mod, PMeco_depth = PMeco_depth)
  })
}

multiMod.fit.plot.fx <- function(fit1, fit1.name, fit2, fit2.name, fit3 = NULL, fit3.name = NULL) {
  lapply(seq_along(fit1), function(i) {
    PMeco_depth <- names(fit1)[i]
    con.df <- con.df.fx(PMeco_depth)
    plot.df <- rbind(fit1[[i]],
                     fit2[[i]],
                     fit3[[i]])
    plot.df$Model <- factor(c(rep(fit1.name, nrow(fit1[[i]])),
                              rep(fit2.name, nrow(fit2[[i]])),
                              rep(fit3.name, nrow(fit3[[i]]))),
                            levels = c(fit1.name, fit2.name, fit3.name))
    return(plot.df %>%
             filter(pool == "bulkC" | pool == "respiration" | pool == "atm") %>%
             ggplot(., aes(years, d14C, color = pool)) +
             geom_path(aes(linetype = Model)) +
             geom_point(data = con.df, aes(Year, d14c, color = pool), size = 3) +
             scale_color_manual(
               name = "Model pool",
               values = c("atm" = 8,
                          "bulkC" = "black",
                          "respiration" = "#e47b1f")) +
             scale_x_continuous(limits = c(1950, 2022)) +
             ggtitle(PMeco_depth) +
             xlab("Year") +
             ylab(expression(''*Delta*''^14*'C (‰)')) +
             theme_bw() +
             theme(panel.grid = element_blank()))
  })
}
```

```{r plot-modFit-curves}
# all 1p mods
plot.ls.1p <- singleMod.fit.plot.fx(mod.fitted.1p, "1p")

# all 2pp mods
plot.ls.2pp <- singleMod.fit.plot.fx(mod.fitted.2pp, "2pp")

# all 2ps mods
plot.ls.2ps <- singleMod.fit.plot.fx(mod.fitted.2ps, "2ps")

# plot both fits
multiMod.fit.plot.fx(mod.fitted.2pp[ix.10], "2pp", mod.fitted.2ps[ix.10], "2ps", mod.fitted.1p, "1p")

# plot initial vs. optimized fit
multiMod.fit.plot.fx(
  mod.fitted.2pp[ix.10], "2pp, fitted", 
  lapply(pars.i.2pp[ix.10], function(x) 
    modFun(x, "2pp", 1, out = "", verbose = FALSE)), "2pp, initial")

multiMod.fit.plot.fx(
  mod.fitted.2ps[ix.10], "2ps, fitted", 
  lapply(pars.i.2ps[ix.10], function(x) 
    modFun(x, "2ps", 1, out = "", verbose = FALSE)), "2ps, initial")
```

```{r test-alt-mod-funs}
mod.sens.fits.2pp <- mod.fits.fx2(mod = "2pp",
                                 pars = pars.i.2pp,
                                 In = in.1.ls,
                                 sub = ix.10,
                                 upper = c(1, 1, 1),
                                 lower = c(0, 0, 0),
                                 cost = "14C")
names(mod.sens.fits.2pp) <- names(pars.i.2pp)[ix.10]
```